---
title: "SuperBowl Predictions"
author: "Andhy Gomez"
date: "1/27/2021"
output: html_document
---

# Load libraries
```{r}
# Load the tidyverse
library(tidyverse)
library(readxl)
```

```{r}
superbowl <- read.csv("superbowl.csv")

superbowl$Date = as.Date(superbowl$Date, "%b %d %Y")

superbowl = superbowl[order(superbowl$Date),]

head(superbowl)
```

```{r}
superbowl$Winner <- as.factor(superbowl$Winner)
superbowl$Loser <- as.factor(superbowl$Loser)
```

```{r}
summary(superbowl)
```


```{r}
ggplot(data = superbowl, aes(Winner)) + 
  stat_count(show.legend = FALSE, aes(color = Winner, fill = Winner)) +
  labs(title = "Superbowl Wins",
       subtitle = "1967 - 2020",
       x = "Team",
       y = "Wins",
       caption = "Source: Kaggle") +
  coord_flip() +
  theme(panel.background = element_blank())
```
```{r}
ggplot(data = superbowl, aes(Loser)) + 
  stat_count(show.legend = FALSE, aes(color = Loser, fill = Loser)) +
  labs(title = "Superbowl Losses",
       subtitle = "1967 - 2020",
       x = "Team",
       y = "Losses",
       caption = "Source: Kaggle") +
  coord_flip() +
  theme(panel.background = element_blank())
```

Objectively speaking, the Kansas City Chiefs have both won more and participated in more total superbowls than the Tampa Bay Bucaneers, That extra experience
as an organization could lead to the Chiefs taking home the championship this year.


```{r}
ggplot(data = superbowl, aes(x = Date)) +
  geom_line(aes(y = Winner.Pts), color = "royalblue3") +
  geom_line(linetype = "dashed", aes(y = Loser.Pts), color = "firebrick4") +
  theme(panel.background = element_blank()) +
  theme( axis.line = element_line(colour = "black", 
                      size = 1, linetype = "solid")) +
  labs(title = "Winners Vs. Losers",
       subtitle = "Superbowl Performances since 1967",
       x = "",
       y = "Points Scored",
       caption = "Blue = Winners
                  Red = Losers")
```
```{r}
mean(superbowl$Winner.Pts)
mean(superbowl$Loser.Pts)
```

On average, the winning team will score 30. The losing team on the other hand generally scores around 16. It stands to reason that the team who has a higher 
score average over the season will likely win. Once again, the Kansas City Chiefs seem to be the favorites.


```{r}
superbowlmvps <- read.csv("mvpsb.csv")
superbowlmvps$POSITION <- as.factor(superbowlmvps$POSITION)
head(superbowlmvps)
```

```{r}
ggplot(data.frame(superbowlmvps), aes(x = POSITION)) +
  geom_bar(show.legend = FALSE, aes(fill = TEAM)) +
  theme(panel.background = element_blank()) +
  theme( axis.line = element_line(colour = "black", 
                      size = 1, linetype = "solid")) +
  labs(title = "Superbowl MVP Winners by Position",
       subtitle = "1967 - 2020",
       x = "Position",
       y = "MVP Count")
  
```

Since 1967, it can be clearly seen that a staggering number of times the Quarterback is declared the MVP of the superbowl, sitting at upwards of 30 QB 
MVP's with the next most being wide receivers which are under 10. Therefore it should not be a stretch to say the team with the better QB will win. 
Although the Buccaneers have a very prolific and historically good QB in Tom Brady, the Chiefs QB Patrick Mahomes is nothing short of a generational talent 
himself, having even won and took home the Championship and MVP honors the prior season. All in all, my favorites to win this years superbowl are the Kansas 
City Chiefs.
